---
title: "rgph2024tn -- Package Validation Report"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
.pass <- 0L; .fail <- 0L
check <- function(desc, expr) {
  ok <- tryCatch({ stopifnot(isTRUE(eval(expr, parent.frame()))); TRUE }, error = function(e) FALSE)
  if (ok) { .pass <<- .pass + 1L; cat("  PASS", desc, "\n") }
  else    { .fail <<- .fail + 1L; cat("  FAIL", desc, "-- FAILED\n") }
}
```

# 1. Install and load

To install from the tarball, place `rgph2024tn-R-0.1.0.tar.gz` next to this
Rmd and knit. Or install from the package directory with `devtools::install()`.

```{r install}
# Try tarball first, fall back to devtools::install()
tarball <- "rgph2024tn-R-0.1.0.tar.gz"
if (file.exists(tarball)) {
  install.packages(tarball, repos = NULL, type = "source")
} else if (requireNamespace("devtools", quietly = TRUE)) {
  devtools::install(".", upgrade = "never", quiet = TRUE)
} else {
  stop("Place rgph2024tn-R-0.1.0.tar.gz here or install devtools.")
}
library(rgph2024tn)
cat("Loaded: rgph2024tn", as.character(packageVersion("rgph2024tn")), "\n")
```

# 2. Metadata

```{r metadata}
d <- packageDescription("rgph2024tn")
check("Package name", d$Package == "rgph2024tn")
check("Version 0.1.0", d$Version == "0.1.0")
check("License MIT", grepl("MIT", d$License))
check("Depends R >= 4.0", grepl("4\\.0", d$Depends))
check("Imports only utils", trimws(d$Imports) == "utils")
check("No heavy deps in Imports", !any(grepl("dplyr|readr|tibble|arrow|rlang", d$Imports)))
check("CITATION exists", file.exists(system.file("CITATION", package = "rgph2024tn")))
check("At least 2 citation entries", length(citation("rgph2024tn")) >= 2)
```

# 3. Delegations (bundled)

```{r delegations}
del <- load_delegations()
check("279 rows", nrow(del) == 279)
check(">= 900 columns", ncol(del) >= 900)
check("First col = code_gouvernorat", names(del)[1] == "code_gouvernorat")
check("Has delegation_fr", "delegation_fr" %in% names(del))
check("Has lat", "lat" %in% names(del))
check("Has lon", "lon" %in% names(del))
check("Has region", "region" %in% names(del))
check("Has code_delegation", "code_delegation" %in% names(del))
check("No NA coords", !any(is.na(del$lat)) && !any(is.na(del$lon)))
```

# 4. Gouvernorats (bundled)

```{r gouvernorats}
gov <- load_gouvernorats()
check("24 rows", nrow(gov) == 24)
check(">= 1000 cols", ncol(gov) >= 1000)
check("First col clean", names(gov)[1] == "code_gouvernorat")
check("Has gouvernorat_fr", "gouvernorat_fr" %in% names(gov))
check("No NA coords", !any(is.na(gov$lat)) && !any(is.na(gov$lon)))
```

# 5. National (bundled)

```{r national}
nat <- load_national()
check("1 row", nrow(nat) == 1)
check(">= 1000 cols", ncol(nat) >= 1000)
```

# 6. Secteurs (requires download)

```{r secteurs}
check("load_secteurs() errors with helpful message",
      tryCatch({ load_secteurs(); FALSE },
               error = function(e) grepl("download_secteurs", e$message)))
```

# 7. Caching

```{r caching}
del2 <- load_delegations()
check("Caching: identical object returned", identical(del, del2))
```

# 8. Accessors

```{r accessors}
tunis <- get_gouvernorat(11)
check("get_gouvernorat(11) = 1 row", nrow(tunis) == 1)
check("get_gouvernorat(11) = Tunis", tunis$gouvernorat_fr == "Tunis")

sfax <- get_gouvernorat(name = "Sfax")
check("get_gouvernorat(name='Sfax') = 1 row", nrow(sfax) == 1)

empty <- get_gouvernorat(99)
check("get_gouvernorat(99) = 0 rows", nrow(empty) == 0)

empty2 <- get_gouvernorat(name = "NONEXISTENT_XYZZY")
check("No match by name = 0 rows", nrow(empty2) == 0)

carth <- get_delegation(1151)
check("get_delegation(1151) = 1 row", nrow(carth) == 1)
check("get_delegation(1151) contains Carthage",
      grepl("Carthage", carth$delegation_fr))

kass <- get_delegation(name = "Kasserine")
check("get_delegation(name='Kasserine') >= 1 row", nrow(kass) >= 1)

co <- get_region("Centre Ouest")
check("get_region('Centre Ouest') has data", nrow(co) > 0)
check("All rows are Centre Ouest", all(co$region == "Centre Ouest"))
```

# 9. Search and codebook

```{r search}
inds <- list_indicators("delegations")
check("list_indicators returns character", is.character(inds))
check("Contains 'emploi'", "emploi" %in% inds)
check("Contains 'biens_equipement'", "biens_equipement" %in% inds)
check("Contains 'migration_interne'", "migration_interne" %in% inds)

cols <- search_indicator("delegations", "chomage")
check("search 'chomage' finds columns", length(cols) > 0)
check("All contain chomage", all(grepl("chomage", cols, ignore.case = TRUE)))

cnam_cols <- search_indicator("delegations", "CNAM")
check("search 'CNAM' finds columns", length(cnam_cols) > 0)

cb <- codebook()
check("codebook returns data.frame", is.data.frame(cb))
check("codebook >= 46 rows", nrow(cb) >= 46)
check("codebook has 4 columns", ncol(cb) == 4)
check("Columns: indicator, domain, description, deepest_level",
      identical(names(cb), c("indicator", "domain", "description", "deepest_level")))

cb_ed <- codebook("education")
check("codebook('education') filters", all(cb_ed$domain == "education"))
check("At least 5 education indicators", nrow(cb_ed) >= 5)

check("8 domains", length(unique(cb$domain)) == 8)
expected_domains <- c("education", "demography", "housing", "infrastructure",
                      "assets", "health", "disability", "employment")
check("All domain names correct", setequal(unique(cb$domain), expected_domains))
```

# 10. Gini coefficient

```{r gini}
check("gini(ones) = 0", gini(rep(1, 100)) == 0)
check("gini(extreme) > 0.98", gini(c(rep(0, 99), 100)) > 0.98)
check("gini(single value) = NA", is.na(gini(5)))
check("gini(empty) = NA", is.na(gini(numeric(0))))
check("gini(all zeros) = NA", is.na(gini(rep(0, 10))))
check("gini(all NA) = NA", is.na(gini(c(NA, NA, NA))))
check("gini handles NA in data", !is.na(gini(c(1, 2, NA, 4, 5))))
check("Gas Gini > 0.5", gini(del[["raccordement.total.Gaz naturel"]]) > 0.5)
check("Elec Gini < 0.05", gini(del[["raccordement.total.STEG"]]) < 0.05)
```

# 11. Theil decomposition

```{r theil}
r <- theil_decompose(del[["raccordement.total.SONEDE"]], del[["region"]])
check("Returns list", is.list(r))
check("Has 5 components",
      all(c("total","between","within","pct_between","group_detail") %in% names(r)))
check("Additivity: total = between + within",
      abs(r$total - r$between - r$within) < 0.01)
check("pct_between in (0, 100)", r$pct_between > 0 && r$pct_between < 100)
check("group_detail is data.frame", is.data.frame(r$group_detail))
check("6 groups (regions)", nrow(r$group_detail) == 6)
check("group_detail has correct columns",
      all(c("group","n","mean","theil","weight") %in% names(r$group_detail)))
```

# 12. Geographic coordinates

```{r coordinates}
check("Del lat in [30, 38]", all(del$lat > 30 & del$lat < 38))
check("Del lon in [7, 12]", all(del$lon > 7 & del$lon < 12))
check("Gov lat in [30, 38]", all(gov$lat > 30 & gov$lat < 38))
check("Gov lon in [7, 12]", all(gov$lon > 7 & gov$lon < 12))
```

# 13. Administrative hierarchy

```{r hierarchy}
gov_codes <- unique(gov$code_gouvernorat)
del_gov_codes <- unique(del$code_gouvernorat)
check("Del gov codes subset of gov codes", all(del_gov_codes %in% gov_codes))

expected_regions <- c("Nord Est", "Nord Ouest", "Centre Est",
                      "Centre Ouest", "Sud Est", "Sud Ouest")
check("All 6 regions present", setequal(unique(del$region), expected_regions))
check("No NA regions", !any(is.na(del$region)))
```

# 14. Key indicator columns exist with data

```{r key-indicators}
key_cols <- c(
  "emploi.taux_chomage.Masculin",
  "emploi.taux_emploi.Feminin",
  "emploi.taux_participation.Feminin",
  "raccordement.total.SONEDE",
  "raccordement.total.STEG",
  "raccordement.total.Gaz naturel",
  "biens_equipement.total.Voiture",
  "migration_interne.sortants",
  "migration_interne.entrants",
  "migration_interne.solde_migratoire",
  "migration_externe.emigrants",
  "affiliation_couverture_sociale.all.Total",
  "couverture_sanitaire_type.all.CNAM.Total"
)

for (col in key_cols) {
  check(paste0("'", col, "' has data"),
        col %in% names(del) && sum(!is.na(del[[col]])) > 200)
}
```

# 15. Value ranges

```{r value-ranges}
prop_cols <- c("raccordement.total.SONEDE", "raccordement.total.STEG",
               "raccordement.total.Gaz naturel", "biens_equipement.total.Voiture",
               "affiliation_couverture_sociale.all.Total")
for (col in prop_cols) {
  v <- del[[col]]; v <- v[!is.na(v)]
  check(paste0(col, " in [0, 1]"), all(v >= 0 & v <= 1.001))
}

m <- del[["migration_interne.solde_migratoire"]]
m <- m[!is.na(m)]
check("Net migration has positive and negative values",
      any(m > 0) && any(m < 0))
```

# 16. Spot checks

```{r spot-checks}
tunis_u <- gov[gov$code_gouvernorat == 11, "emploi.taux_chomage.Masculin"]
check("Tunis male unemployment < 20%", !is.na(tunis_u) && tunis_u < 0.20)

carth_lat <- del[del$code_delegation == 1151, "lat"]
carth_lon <- del[del$code_delegation == 1151, "lon"]
check("Carthage lat near 36.85", abs(carth_lat - 36.85) < 0.1)
check("Carthage lon near 10.32", abs(carth_lon - 10.32) < 0.1)

gas <- del[["raccordement.total.Gaz naturel"]]
gas <- gas[!is.na(gas)]
check(">50% delegations have <5% gas", mean(gas < 0.05) > 0.5)
```

# Summary

```{r summary, echo=FALSE}
cat("\n")
cat("======================================\n")
cat(sprintf("  Passed: %d\n", .pass))
cat(sprintf("  Failed: %d\n", .fail))
cat("======================================\n")
if (.fail == 0) {
  cat("\n  PASS ALL TESTS PASSED\n")
  cat("  Ready for: R CMD build && R CMD check --as-cran\n")
} else {
  cat(sprintf("\n  FAIL %d TESTS FAILED\n", .fail))
}
```

---

*Generated `r Sys.time()` on `r R.version.string`*
